Supplementary Information

### 2020-10-22 writing/transect_supplement.Rmd
radical reorganization of light plan and scripts directory.  everything works again.

datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccccccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
if (!exists("percent", mode = "function")) source(here::here("scripts", "output_functions.R"))

@@

**Supplementary file 3**: ML tree of all long metabarcoding ASVs (5.8S--ITS2--LSU) assigned to kingdom Fungi.
Thick branches represent rapid bootstrap values >90%.
Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, and Unite, Warcup, and RDP reference databases. 
Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
Taxon names in red were reconstructed as polyphyletic.
In some cases, this may be due to a poorly identified sequence "splitting" the group (e.g., the single *Aspergillus* sequence nested in *Monascus*, Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).

@@

## Sampling

@@

## DNA extraction, amplification, and sequencing

@@

## Denoising and clustering

@@

## Phylogenetic inference and taxonomy assignment

@@

The phylogenetic tree was rooted based on sequences which were confidently assigned to kingdom based on the primary taxonomic assignments.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
Assignments were considered as reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
Two sequences belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(read(sure_kingdoms), taxon) %>% glue::glue("{n} assigned to {taxon}, ", last = "and ")`.
Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
The kingdom Fungi was then identified as the minimal clade containing all ASVs which were confidently identified  as kingdom Fungi.
ASVs falling outside this clade, including the excluded *Tulasnella* sequences, were not included in downstream fungal community analysis.

### Refining taxonomic assignments with a phylogenetic tree: PHYLOTAX

@@

## Effect of sequencing strategy on recovered community

@@

## Spatiotemporal analysis

@@

```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table")}
plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
points(
  readd(buffer_asv_pcoa),
  select = with(
    as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
    buffer == "Xpedition" & year == "2015" 
  ),
  col = "green",
  pch = 1
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "Xpedition" & year == "2016"
  ),
  col = "blue",
  pch = 2
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))),
                buffer == "LifeGuard" & year == "2016"
  ),
  col = "red",
  pch = 3
)
legend(
  "bottomright",
  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
  col = c("green", "blue", "red"),
  pch = 1:3
)
# legend(
#   "topright",
#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
#     unique() %>%
#     sort(),
#   lty = 1:3
# )
# vegan::ordiellipse(
#   readd(buffer_asv_pcoa),
#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
#   lty = 1:3
# )
```

@@

(ref:alpha-compare-OTU) Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies. All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.  Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction. Blue line is a total least squares fit, with the given slope (and 95% confidence interval) and R^2^ value. Dashed diagonal line indicates a slope of 1.

@@

```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
readd(big_seq_table_ITS2) %>%
  colSums() %>%
  enframe(name = "ITS2", value = "nreads") %>%
  mutate(length = nchar(ITS2)) %>%
  filter(length <= 140) %>%
  mutate_at("ITS2", ~tzara::seqhash(chartr("T", "U", .))) %>%
  semi_join(readd(taxon_reads), ., by = c("label" = "ITS2")) %>%
  group_by(tech, amplicon, Algorithm, reference) %>%
  mutate(reads = reads/sum(reads)) %>%
  ungroup() %>%
  # group_by(ITS2) %>%
  # filter(!"Short" %in% amplicon) %>%
  taxon_plot(kingdom, y = reads, datasets = datasets)
```

@@

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
  readd(taxon_reads) %>%
  filter(ifelse(Algorithm == "Consensus", region == "All", TRUE)) %>%
  mutate_at("reference", replace_na, "All") %>%
  taxon_plot(ECM, kingdom == "Fungi", datasets = datasets) +
  theme(legend.position = "bottom", legend.title = element_blank())
```

>>>>

datasets <- read_csv(here("config/datasets.csv"), col_types = "cccccccicccccccccccicc")
regions <- read_csv(here("config/regions.csv"), col_types = "cccciiiiic")
if (!exists("percent", mode = "function")) source(here::here("scripts", "functions_output.R"))

@@

**Supplementary file 3**: ML tree of all long metabarcoding ASVs (ITS1--5.8S--ITS2--LSU) assigned to kingdom Fungi.
Thick branches represent rapid bootstrap values >90%.
Taxonomic assignment by PHYLOTAX, based on assignments using RDPC, SINTAX, and IDTAXA, and Unite, Warcup, and RDP reference databases. 
Taxon names in blue were not fully resolved with PHYLOTAX, and are given as a list of alternative assignments, along with the number of methods which gave each assignment (max 9).
Taxon names in red were reconstructed as polyphyletic.
In some cases, this may be due to a poorly identified sequence "splitting" the group (e.g., the single *Aspergillus* sequence nested in *Monascus*, Aspergillaceae and Trichocomaceae sequences nested in Onygenales), but in other cases, the tree has clearly failed to reconstruct groups which are generally considered to be monophyletic based on multi-gene trees (e.g., Dothidiomycetes, Cantharellales).

@@

## Sampling {#Sampling-supplement}

@@

## DNA extraction, amplification, and sequencing {#DNA-supplement}

@@

## Denoising and clustering {#denoising-supplement}

@@

## Phylogenetic inference and taxonomy assignment {#phylogeny-supplement}

@@

The phylogenetic tree was rooted based on sequences which were confidently assigned to kingdom based on the primary taxonomic assignments.
Assignments based on Warcup were not used at this step because non-Fungi are not included in the dataset, and thus it cannot be used to directly assign kingdoms.
Assignments were considered as reliable if at least five of the six remaining primary assignment methods (2 databases × 3 algorithms) resulted in the same kingdom-level assignment, with no conflicting assignments.
Two sequences belonging to *Tulasnella*, which is known to have an accelerated rate of molecular evolution in the rDNA and often forms long, unstable branches in phylogenies [@moncalvo2006], were excluded at this step.
This resulted in `r nrow(readd(sure_kingdoms))` confident kingdom assignments, with `r dplyr::count(readd(sure_kingdoms), taxon) %>% glue::glue_data("{n} assigned to {taxon}, ") %>% glue::glue_collapse(last = "and ")`.
Inspection of the tree verified that these four kingdoms were reciprocally monophyletic, and the clade containing Alveolata and Viridiplantae was used as an outgroup to root the tree.
The kingdom Fungi was then identified as the minimal clade containing all ASVs which were confidently identified  as kingdom Fungi.
ASVs falling outside this clade, including the excluded *Tulasnella* sequences, were not included in downstream fungal community analysis.

### Refining taxonomic assignments with a phylogenetic tree: `PHYLOTAX`

@@

## Effect of sequencing strategy on recovered community {#sequencing-strategy-supplement}

@@

## Spatiotemporal analysis {#spatiotemporal-supplement}

@@

```{r buffer-pcoa, fig.cap="(ref:buffer-pcoa-s), (ref:buffer-pcoa)", fig.scap = "(ref:buffer-pcoa-s)", cache = TRUE, extra.cache = gethash("buffer_asv_pcoa", "buffer_asv_table")}
plot(readd(buffer_asv_pcoa), type = "none", xlab = "PCoA1", ylab = "PCoA2")
points(
  readd(buffer_asv_pcoa),
  select = with(
    as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
    buffer == "Xpedition" & year == "2015" 
  ),
  col = "green",
  pch = 1
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
                buffer == "Xpedition" & year == "2016"
  ),
  col = "blue",
  pch = 2
)
points(
  readd(buffer_asv_pcoa),
  select = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_physeq))),
                buffer == "LifeGuard" & year == "2016"
  ),
  col = "red",
  pch = 3
)
legend(
  "bottomright",
  legend = c("Xpedition 2015", "Xpedition 2016", "LifeGuard 2016"),
  col = c("green", "blue", "red"),
  pch = 1:3
)
# legend(
#   "topright",
#   legend = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)) %>%
#     unique() %>%
#     sort(),
#   lty = 1:3
# )
# vegan::ordiellipse(
#   readd(buffer_asv_pcoa),
#   groups = with(as_tibble(phyloseq::sample_data(readd(buffer_asv_table))), paste(tech, amplicon)),
#   lty = 1:3
# )
```

@@

(ref:alpha-compare-OTU) Each point represents the richness of one or two pooled samples, as determined by two different sequencing strategies. All values represent the average of 100 replicate rarefactions with a sample depth of 100 reads.  Because samples in the same well on different plates could not be demultiplexed in the Ion Torent dataset, these samples were also bioinformatically pooled in the other datasets prior to rarefaction. Blue lines are total least squares fit with 95% confidence interval, with the given slope (and 95% confidence interval) and R^2^ value. Dashed diagonal line indicates ideal slope of 1.

@@

```{r short-its, fig.cap = "(ref:short-its-s). (ref:short-its)", fig.scap = "(ref:short-its-s)", fig.width=7, cache=TRUE, extra.cache=gethash("big_seq_table_ITS2", "taxon_reads"), more.cache = datasets}
readd(taxon_plot_short_kingdom)
```

@@

```{r ecm, fig.cap = "(ref:ecm-s). (ref:ecm)", fig.scap = "(ref:ecm-s)", fig.width = 7, cache=TRUE, extra.cache=gethash("taxon_reads"), more.cache = datasets}
readd(taxon_plot_ecm)
```
